---
title: "Games & Puzzles"  
author: "Bernardo Lares"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Games & Puzzles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)

# Enable ANSI color output in HTML vignettes
options(crayon.enabled = TRUE)
library(fansi)
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks,
  which = c("output", "message", "error")
)
```

## Introduction

Beyond data analysis and machine learning, `lares` includes a collection of fun game solvers and helpers! Challenge yourself with Wordle, Scrabble, Sudoku, mazes...

```{r}
library(lares)
```

## Wordle

Play or solve Wordle puzzles with validation, hints, and simulations.

### Basic Wordle Validation

```{r, eval=TRUE}
# Check your guess against a word
wordle_check("OPENS", "ABBEY")
wordle_check("BABES", "ABBEY")
wordle_check("ABBEY", "ABBEY")
```

### Get Wordle Hints

Use `scrabble_words()` to find possible words:

```{r}
# After OPENS: O is not in word, P/E/N/S are not in positions 2/3/4/5
hints <- scrabble_words(
  tiles = "abcdefghijklmrtuvwxyz", # Available letters
  exclude_here = list("2" = "p", "3" = "e", "4" = "n", "5" = "s"),
  force_exclude = c("o"),
  force_n = 5, # 5-letter words
  language = "en"
)
head(hints, 10)
```

### Wordle Simulation

Simulate solving a Wordle puzzle:

```{r}
# Simulate solving with different starting words
simulation <- wordle_simulation(
  input = "SAINT",
  word = "ABBEY",
  seed = 123
)
print(simulation)
```

### Wordle Dictionary

Access word lists for different languages:

```{r eval=FALSE}
# Get English 5-letter words
en_words <- wordle_dictionary("en")
head(en_words, 20)

# Spanish words
es_words <- wordle_dictionary("es")
```

## Scrabble

Maximize your Scrabble score with word finders and calculators!

### Find Highest-Scoring Words

```{r}
# Find best words from your tiles
scrabble_words(
  tiles = "aeiourtn",
  force_max = 8, # Max 8 letters
  language = "en",
  scores = "en"
)
```

### With Board Constraints

```{r}
# Must contain specific letters
scrabble_words(
  tiles = "bernardo",
  force_str = "arn",
  force_max = 7,
  language = "en"
)
```

### Calculate Word Scores

```{r}
# Get point values for each letter
en_scores <- scrabble_points("en")
print(en_scores)

# Calculate scores for words
words <- c("QUEEN", "QUIZ", "HELLO")
scrabble_score(words, en_scores)
```

### Multi-Language Support

```{r eval=FALSE}
# Spanish Scrabble
scrabble_words(
  tiles = "españa",
  language = "es",
  scores = "es"
)

# French Scrabble
scrabble_words(
  tiles = "bonjour",
  language = "fr",
  scores = "fr"
)
```

## Sudoku

Solve Sudoku puzzles automatically!

### Simple Sudoku

```{r}
# Easy puzzle (0 represents empty cells)
trivial <- matrix(c(
  0, 9, 0, 7, 0, 0, 8, 6, 0,
  0, 3, 1, 0, 0, 5, 0, 2, 0,
  8, 0, 6, 0, 0, 0, 0, 0, 0,
  0, 0, 7, 0, 5, 0, 0, 0, 6,
  0, 0, 0, 3, 0, 7, 0, 0, 0,
  5, 0, 0, 0, 1, 0, 7, 0, 0,
  0, 0, 0, 0, 0, 0, 1, 0, 9,
  0, 2, 0, 6, 0, 0, 3, 5, 0,
  0, 5, 4, 0, 0, 8, 0, 7, 0
), nrow = 9, byrow = TRUE)

solution <- sudoku_solver(trivial)
print(solution)
```

### Complex Sudoku

```{r eval=FALSE}
# Harder puzzle
difficult <- matrix(c(
  5, 3, 0, 0, 7, 0, 0, 0, 0,
  6, 0, 0, 1, 9, 5, 0, 0, 0,
  0, 9, 8, 0, 0, 0, 0, 6, 0,
  8, 0, 0, 0, 6, 0, 0, 0, 3,
  4, 0, 0, 8, 0, 3, 0, 0, 1,
  7, 0, 0, 0, 2, 0, 0, 0, 6,
  0, 6, 0, 0, 0, 0, 2, 8, 0,
  0, 0, 0, 4, 1, 9, 0, 0, 5,
  0, 0, 0, 0, 8, 0, 0, 7, 9
), nrow = 9, byrow = TRUE)

sudoku_solver(difficult)
```

## Maze Solver

Solve mazes using depth-first search algorithms!

### Basic Maze

```{r}
# Create a simple maze (0 = path, 1 = wall)
simple_maze <- matrix(c(
  0, 1, 0, 0, 0,
  0, 1, 0, 1, 0,
  0, 0, 0, 1, 0,
  1, 1, 0, 0, 0,
  0, 0, 0, 1, 0
), nrow = 5, byrow = TRUE)

solution <- maze_solve(
  simple_maze,
  start = c(1, 1),
  end = c(5, 5)
)
print(solution)
```

### Micromouse Competition Maze

```{r eval=FALSE}
# Classic Micromouse-style maze
micromouse <- matrix(c(
  1, 1, 1, 1, 1, 1, 1, 1,
  1, 0, 0, 0, 0, 0, 0, 1,
  1, 0, 1, 1, 1, 1, 0, 1,
  1, 0, 1, 0, 0, 0, 0, 1,
  1, 0, 1, 0, 1, 1, 0, 1,
  1, 0, 0, 0, 0, 1, 0, 1,
  1, 0, 1, 1, 0, 0, 0, 1,
  1, 1, 1, 1, 1, 1, 1, 1
), nrow = 8, byrow = TRUE)

maze_solve(
  micromouse,
  start = c(2, 2),
  end = c(7, 7),
  diagonal = FALSE
)
```

### Advanced Options

```{r eval=FALSE}
# With diagonal movement and aiming toward goal
maze_solve(
  micromouse,
  start = c(2, 2),
  end = c(7, 7),
  diagonal = TRUE, # Allow diagonal moves
  aim = TRUE, # Prefer directions toward goal
  inertia = TRUE, # Prefer continuing in same direction
  timeout = 5 # Max 5 seconds
)
```

## Combining Games & Data Science

Use game functions for:

**Text Analysis:**
```{r eval=FALSE}
# Find anagrams in your dataset
words <- c("listen", "silent", "hello")
scrabble_words(tiles = "listen", language = "en")
```

**Algorithm Teaching:**
```{r eval=FALSE}
# Demonstrate recursion with maze solving
maze_solve(simple_maze, start = c(1, 1), end = c(5, 5))
```

**Pattern Recognition:**
```{r eval=FALSE}
# Wordle simulations for optimal starting words
seeds <- 1:100
results <- lapply(seeds, function(s) {
  wordle_simulation("SAINT", "ABBEY", seed = s, quiet = TRUE)
})
```

## Further Reading

### Package Resources
- **Package documentation:** [https://laresbernardo.github.io/lares/](https://laresbernardo.github.io/lares/)
- **GitHub repository:** [https://github.com/laresbernardo/lares](https://github.com/laresbernardo/lares)
- **Report issues:** [https://github.com/laresbernardo/lares/issues](https://github.com/laresbernardo/lares/issues)

### Game References
- **Wordle:** The viral word game by Josh Wardle
- **Sudoku:** Japanese number puzzle
- **Scrabble:** Classic word board game
- **Micromouse:** Autonomous maze-solving competition

## Next Steps

- Explore data analysis features (see Data Wrangling vignette)
- Learn machine learning (see Machine Learning vignette)
- Try API integrations (see API Integrations vignette)
- Check game function documentation: `?wordle_check`, `?scrabble_words`, `?sudoku_solver`, `?maze_solve`
