---
title: "Robyn Enhancers"
author: "Bernardo Lares"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Robyn enhancers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.width = 7,
    fig.height = 5,
    warning = FALSE,
    message = FALSE,
    cache = TRUE
)
library(lares)
library(dplyr)
library(ggplot2)
```

## Introduction

The `lares` package provides a suite of auxiliary functions to enhance your experience with [Robyn](https://github.com/facebookexperimental/Robyn), a Marketing Mix Modeling (MMM) package developed by Meta's Marketing Science Team. These functions help with hyperparameter building, model selection, performance analysis, and cross-model budget allocation.

## Hyperparameter Builder

The function `robyn_hypsbuilder()` automates the creation of hyperparameter lists for Robyn. It's particularly useful when you have many media channels and want to set reasonable default ranges.

```{r eval=TRUE}
# Define your input variables
paid_media_spends <- c("tv_spend", "digital_spend")
organic_vars <- c("newsletter")

# Generate hyperparameters
hyperparameters <- robyn_hypsbuilder(
    channels = c(paid_media_spends, organic_vars),
    media_type = "default",
    adstock = "weibull_pdf", # or "weibull_cdf", "weibull_pdf"
    date_type = "weekly",
    lagged = FALSE
)

# The output is a list ready to be passed to robyn_inputs()
print(hyperparameters)
```

## Model Selection

The function `robyn_modelselector()` provides a visual and metric-based approach to select the best models from a Robyn run. It ranks models based on their performance metrics and helps you identify the best candidates to evaluate. For a deeper dive into this topic, check out this [Medium post](https://medium.com/@laresbernardo/select-the-right-mmm-candidate-based-on-your-specific-criteria-and-business-knowledge-1f583c3cb97a).

```{r eval=FALSE}
# Assuming you have a Robyn OutputCollect object
# OutputCollect <- robyn_run(...)

# Select best models
best_models <- robyn_modelselector(
    InputCollect = InputCollect,
    OutputCollect = OutputCollect,
    metrics = c(
        "rsq_train", "performance", "potential_improvement", "non_zeroes",
        "incluster_models", "cluster_sd", "certainty", "baseline_dist"
    ),
    wt = c(2, 0.1, 0, 1, 0.1, 0, 1.5, 0)
)

# View the results
print(best_models)

# Plot the clusters
plot(best_models)
```

## Model Performance

The function `robyn_performance()` calculates detailed performance metrics for a specific model or a set of models. It breaks down contribution, ROAS or CPA for each channel, etc.

```{r eval=FALSE}
# Get performance for a specific model and date range
performance <- robyn_performance(
    InputCollect = InputCollect,
    OutputCollect = OutputCollect,
    select_model = "1_123_4",
    # Calculate for specific date range
    start_date = NULL, end_date = NULL,
    non_promo = TRUE, # Add non-promo decomposition as well
    marginals = TRUE, # Add mROAS or mCPA for spend levels
    carryovers = FALSE # Add mean percentage of carryover
)

# View performance table
print(performance)

# Plot performance
plot(performance)
```

## Marginal Performance (Experimental)

The function `robyn_marginal()` allows you to explore the marginal returns of your media channels (mROAS or mCPA).

```{r eval=FALSE}
# Calculate marginal response
marginal <- robyn_marginal(
    InputCollect = InputCollect,
    OutputCollect = OutputCollect,
    select_model = "1_123_4",
    metric_name = "tv_spend",
    metric_value = 100000
)
```

## Cross-Model Budget Allocation (Experimental)

If you have multiple Robyn models (e.g., for different countries or brands), `robyn_xmodels()` and `robyn_xchannels()` help you allocate budget across all of them simultaneously.

```{r eval=FALSE}
# Load multiple models
model_paths <- c("model_us.json", "model_uk.json", "model_de.json")
models <- lapply(model_paths, Robyn::robyn_recreate)
names(models) <- c("US", "UK", "DE")

# Allocate budget across models
cross_allocation <- robyn_xmodels(
    models = models,
    total_budget = 5000000,
    scenario = "max_response"
)

# View results
print(cross_allocation)
```

## Conclusion

These functions streamline the post-modeling analysis in Robyn, making it easier to interpret results, select the best models, and make data-driven budget allocation decisions.

## Additional Reading

For more information on Robyn and MMM, check out the following resources:

-   [Robyn GitHub Repository](https://github.com/facebookexperimental/Robyn)
-   [Robyn: Analyst's Guide to MMM](https://facebookexperimental.github.io/Robyn/docs/analysts-guide-to-MMM)
-   [Select the right MMM candidate based on your specific criteria and business knowledge](https://medium.com/@laresbernardo/select-the-right-mmm-candidate-based-on-your-specific-criteria-and-business-knowledge-1f583c3cb97a)
-   [Stop building MMMs without business-backed hypotheses](https://medium.com/@laresbernardo/stop-building-mmms-without-business-backed-hypotheses-0e8a9fa9a375)
-   [Hitting ROAS Target Using Robyn’s Budget Allocator](https://medium.com/@gufengzhou/hitting-roas-target-using-robyns-budget-allocator-274ace3add4f)
-   [The Convergence of Marginal ROAS in the Budget Allocation in Robyn](https://medium.com/@gufengzhou/the-convergence-of-marginal-roas-in-the-budget-allocation-in-robyn-5d407aebf021)
